---
title: "Viz I"
output: github_document
date: "2025-09-26"
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
```

Import the weather data

```{r}
data("weather_df")
```

Making our first plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather_scatterplot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather_scatterplot
  
```

Check that some rows are missing

```{r}
weather_df |> 
  filter(is.na(tmax))
```

## Fancier scatterplots

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name),alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name),alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

Make a more interesting scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(aes(size = prcp),alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

LA #1

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "lm") ## se = FALSE removes error band on graph
```

## Small things

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
 # geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) 

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_smooth(se = FALSE) + 
  geom_point()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() + 
  facet_grid(name ~ .)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density()  
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin), fill = name) +
  geom_violin()
```

Let's make our basic scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  )
```

## Scales

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location") +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```

## Themes

```{r}
ggp_temperature = weather_df |>  
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location") +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```

Update my base plot

```{r}
ggp_temperature +
  theme_minimal()
  theme(legend.position = "bottom") 
```

## Adding data in geoms

```{r}
central_park_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax)) +
  geom_point() +
  geom_line(data = central_park_df)
```

## `patchwork`

Make three plots and combine using patchwork

```{r}
ggp_tmax_tmin =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "bottom")


ggp_prec_density =
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5)

ggp_temp_season = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prec_density)
```

## Data manipulation

Let's make temperature violin plots.

```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = 0.5)
```

```{r}
weather_df |>
  mutate(name = fct_reorder(name, tmax), .na_rm = FALSE) |>  ## put name in order of tmax largest to smallest
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

What about data tidiness?

```{r}
pulse_df = 
  haven::read_sas("public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  )

pulse_df |> 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

```

